State of the Practice for Lattice Boltzmann Method Software

### 2022-03-17 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Addition of Sadeghi et al references, as suggested by Zahra
Addition of comparison to ocean modelling software

A small sample of important applications of LBM include the following: designing
fuel cells \citep{ZhangEtAl2018}, modelling groundwater flow
\citep{AnwarAndSukop2009}, and analyzing the flow of blood in the cardiovascular
system \citep{SadeghiEtAl2020}.  As these examples illustrate, LBM can be used
for tackling problems that impact such areas as environmental policy,
manufacturing, and health and safety.  Given the important applications of LBM,
users of the libraries will be concerned with software qualities like
reliability, robustness, reproducibility, performance, correctness and
verifiability.  Since their time is valuable, developers that create or modify
LBM libraries will have additional quality concerns, including maintainability,
reusability, and understandability.  With the quality considerations in mind,
and the potentially overwhelming number of choices for existing LBM libraries,
the time is right to assess the state of the practice.  The goal of this report
is to analyze the current state of LBM software development and compare it to
the development practices employed for research software in general.  We want to
highlight success stories that LBM developers can share amongst their community,
and with the broader research software community, while at the same time
looking for areas for potential future improvement.

@@

The majority of LBM generated artifacts correspond to general recommendations
from research software developers.  A union of the three columns in
Table~\ref{artifactspresent} mostly corresponds to recommendations made by the
research software community.  Most of the items in Table~\ref{artifactspresent}
are included in guidelines for writing research software, such as the DLR
(German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, xSDK (Extreme-scale Scientific Software Development Kit)
Community Package Policies \citep{SmithAndRoscoe2018}, the EURISE (European
Research Infrastructure Software Engineers') Network Technical Reference
\cite{ThielEtAl2020} and CLARIAH (Common Lab Research Infrastructure for the
Arts and Humanities) Guidelines for Software Quality \citep{vanGompelEtAl2016}.
As an example, LBM developers commonly list the external libraries on which the
software depends, as recommended by \citet{BrettEtAl2021},
\citet{SmithAndRoscoe2018} and \citet{vanGompelEtAl2016}. Although in the rare
column, some LBM developers include an FAQ (Frequently Asked Questions) document
as recommended by \citep{OrvizEtAl2017}, \citep{ThielEtAl2020} and
\citep{vanGompelEtAl2016}. LBM developers also follow the ubiquitous
recommendation to have a README file, although, as for other software, the
content of README files is not standardized \citep{PranaEtAl2018}.

@@

@@

To address technical debt (\ppref{P_TechDebt}), the top LBM developers show how
modularization can be used to design research software for future change.
Although the advice to modularize research software to handle complexity is
common \citep{WilsonEtAl2014, StewartEtAl2017, Storer2017}, specific guidelines
on how to divide the software into modules is less prevalent.  Not every
decomposition is a good design for supporting change, as shown by
\citet{Parnas1972a}.  A design with low cohesion and high coupling \citep[p.\
48]{GhezziEtAl2003} will make change difficult. Especially in research software,
where change is inevitable, designers need to produce a modularization that
supports change.

@@

Future work can be done to address the threats to validity mentioned in
Section~\ref{threats}.  For instance, the measures listed in our measurement
template could be broadened to better capture some qualities. For example,
usability experiments and performance benchmarks could be incorporated into the
assessment.  Section~\ref{threats} also mentions that some pertinent information
on continuous integration and unit testing was not specified in public
artifacts. Adding further questions to the interview guide regarding the
measures that are on the measurement template could reduce instances of
incomplete data. This additional interview data could be analyzed and
incorporated into the AHP ranking, ensuring quality designations more accurately
represent the true quality of the software packages. In the future we could also
expand the interview questions to go beyond reproducibility to also discuss
replicability. Furthermore, in future interviews we would look to clarify the
distinction between usability and understandability
(Section~\ref{softwarequalities}), since the distinction confused some
participants.

@@

TO DO

- should look at ReinerEtAl2022 on Domain of Ocean Modelling software to
contrast to the state of the practice of LBM software

>>>>

A small sample of important applications of LBM include the following: designing
fuel cells \citep{ZhangEtAl2018}, modelling groundwater flow
\citep{AnwarAndSukop2009}, and analyzing the flow of blood in the cardiovascular
system \citep{SadeghiEtAl2022b, SadeghiEtAl2022, SadeghiEtAl2020}.  As these
examples illustrate, LBM can be used for tackling problems that impact such
areas as environmental policy, manufacturing, and health and safety.  Given the
important applications of LBM, users of the libraries will be concerned with
software qualities like reliability, robustness, reproducibility, performance,
correctness and verifiability.  Since their time is valuable, developers that
create or modify LBM libraries will have additional quality concerns, including
maintainability, reusability, and understandability.  With the quality
considerations in mind, and the potentially overwhelming number of choices for
existing LBM libraries, the time is right to assess the state of the practice.
The goal of this report is to analyze the current state of LBM software
development and compare it to the development practices employed for research
software in general.  We want to highlight success stories that LBM developers
can share amongst their community, and with the broader research software
community, while at the same time looking for areas for potential future
improvement.

@@

The majority of LBM generated artifacts correspond to general recommendations
from research software developers.  A union of the three columns in
Table~\ref{artifactspresent} mostly corresponds to recommendations made by the
research software community.  Most of the items in Table~\ref{artifactspresent}
are included in guidelines for writing research software, such as the DLR
(German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, xSDK (Extreme-scale Scientific Software Development Kit)
Community Package Policies \citep{SmithAndRoscoe2018}, the EURISE (European
Research Infrastructure Software Engineers') Network Technical Reference
\cite{ThielEtAl2020} and CLARIAH (Common Lab Research Infrastructure for the
Arts and Humanities) Guidelines for Software Quality \citep{vanGompelEtAl2016}.
As an example, LBM developers commonly list the external libraries on which the
software depends, as recommended by \citet{BrettEtAl2021},
\citet{SmithAndRoscoe2018} and \citet{vanGompelEtAl2016}. Although in the rare
column, some LBM developers include an FAQ (Frequently Asked Questions) document
as recommended by \citep{OrvizEtAl2017}, \citep{ThielEtAl2020} and
\citep{vanGompelEtAl2016}. LBM developers also follow the ubiquitous
recommendation to have a README file, although, as for other software, the
content of README files is not standardized \citep{PranaEtAl2018}.  Compared to
ocean modelling software \citep{JungEtAl2022}, LBM software pays more
attention to artifacts related to testing.

@@

Based on information provided by \citet{JungEtAl2022}, tool utilization for
LBM software has much in common with tool utilization for ocean modelling
software.  Both use tools for editing, compiling, code management, testing,
building and project management.  From the data available, ocean modelling
differs from LBM software in its use of Kanban boards for project management.
LBM software difference from the ocean domain through the use of continuous
integration, and document generation tools.

@@

To address technical debt (\ppref{P_TechDebt}), the top LBM developers show how
modularization can be used to design research software for future change.
Although the advice to modularize research software to handle complexity is
common \citep{WilsonEtAl2014, StewartEtAl2017, Storer2017}, specific guidelines
on how to divide the software into modules is less prevalent.  Not every
decomposition is a good design for supporting change, as shown by
\citet{Parnas1972a}.  A design with low cohesion and high coupling \citep[p.\
48]{GhezziEtAl2003} will make change difficult. Especially in research software,
where change is inevitable, designers need to produce a modularization that
supports change. Ocean modelling software is currently feeling the pain of
not emphasizing modularization in legacy code \citep{Reiner2022}.

@@

Future work can be done to address the threats to validity mentioned in
Section~\ref{threats}.  For instance, the measures listed in our measurement
template could be broadened to better capture some qualities. For example,
usability experiments and performance benchmarks could be incorporated into the
assessment.  Section~\ref{threats} also mentions that some pertinent information
on continuous integration and unit testing was not specified in public
artifacts. Adding further questions to the interview guide regarding the
measures that are on the measurement template could reduce instances of
incomplete data. This additional interview data could be analyzed and
incorporated into the AHP ranking, ensuring quality designations more accurately
represent the true quality of the software packages. In the future we could also
expand the interview questions to go beyond reproducibility to also discuss
replicability. Furthermore, in future interviews we would look to clarify the
distinction between usability and understandability
(Section~\ref{softwarequalities}), since the distinction confused some
participants.  Analysis of future interview data would likely benefit from a
more rigorous process, like a thematic domain analysis \citep{JungEtAl2022}.

@@
